Climate indicators

### 2023-03-08 slitasje.Rmd
convert to continous

```{r}
naturetypes_summary <- naturetypes_summary_import %>%
  rowwise() %>%
  mutate(keepers = sum(c_across(
    c('7TK', '7SE', 'PRTK', 'PRSL'))>0, na.rm=T)) %>%
  filter(keepers > 0) %>%
  select(Nature_type, NiN_mainType, Year)
```

@@

naturligÅpneOmråderILavlandet"] <- "naturligÅpneOmråderUnderSkoggrensa"

Importing and sub-setting the main data file, fix duplicate _hovedøkosystem_, calculate area, split one column in two, make numeric, and select target variables:
```{r, message=F}
naturetypes <- sf::st_read(dsn = path) %>%
  filter(naturtype %in% naturetypes_summary$Nature_type) %>%
  mutate(hovedøkosystem = recode(hovedøkosystem,
                                 "naturligÅpneOmråderUnderSkoggrensa" = "naturligÅpneOmråderILavlandet"),
         area = st_area(.)) %>%
  separate_rows(ninBeskrivelsesvariable, sep=",") %>%
  separate(col = ninBeskrivelsesvariable,
           into = c("NiN_variable_code", "NiN_variable_value"),
           sep = "_",
           remove=F) %>%
  mutate(NiN_variable_value = as.numeric(NiN_variable_value)) %>%
  filter(NiN_variable_code %in% c("7SE", "7TK", "PRTK", "PRSL"))
```

@@

Now I need to create a single row per locality with a new variable which
is a product/function of the four variables "7SE", "7TK", "PRTK" and
"PRSL". I will create a new data table where I calculate the new
variable which I then paste back into the sf object.

```{r}
naturetypes_wide <- pivot_wider(naturetypes,
                                names_from = "NiN_variable_code",
                                values_from = "NiN_variable_value",
                                id_cols = "identifikasjon_lokalId")
naturetypes_wide <- as.data.frame(naturetypes_wide)
head(naturetypes_wide)
```

@@

I want to base this indicator on whichever of the variables have the
highest value (worst condition). This table shows that a locality may
have been assessed with both a PR variable and a NiN variable:

```{r}
table(naturetypes_wide$var_7SE, naturetypes_wide$var_PRTK)
```

Therefore I need to reconcile these scales. The only way I can think of
is to decrease the resolution of the PR variables to the ar4b scale.
This is of course a shame to loose this detail.

A locality will never have values for both 7SE and PRSL

```{r}
table(naturetypes_wide$var_7SE, naturetypes_wide$var_PRSL)
```

nor 7TK and PRTK

```{r}
table(naturetypes_wide$var_7TK, naturetypes_wide$var_PRTK)
```

```{r}
naturetypes_wide$var_7TK <-  ifelse(is.na(naturetypes_wide$var_7TK),
                              ifelse(naturetypes_wide$var_PRTK == 0, 0,
                                ifelse(naturetypes_wide$var_PRTK < 3, 1,
                                  ifelse(naturetypes_wide$var_PRTK < 6, 2, 3))),
                            naturetypes_wide$var_7TK)
```

Checking that I successfully combined classes 1+2, 3+4+5 and 6+7.

```{r}
table(naturetypes_wide$var_7TK, naturetypes_wide$var_PRTK,
      deparse.level = 2)
```

That looks fine. Now doing the same for the other variable PRSL:

```{r}
naturetypes_wide$var_7SE <-  ifelse(is.na(naturetypes_wide$var_7SE),
                              ifelse(naturetypes_wide$var_PRSL == 0, 0,
                                ifelse(naturetypes_wide$var_PRSL < 3, 1,
                                  ifelse(naturetypes_wide$var_PRSL < 6, 2, 3))),
                            naturetypes_wide$var_7SE)
```

Remove the old PR variables.

```{r}
naturetypes_wide <- select(naturetypes_wide,
                           -var_PRTK,
                           -var_PRSL)
```

There are some very few cells with X's. These are NA's.

```{r}
naturetypes_wide$var_7SE[naturetypes_wide$var_7SE == "X"] <- NA
naturetypes_wide$var_7TK[naturetypes_wide$var_7TK == "X"] <- NA

naturetypes_wide$var_7SE <- as.numeric(naturetypes_wide$var_7SE)
naturetypes_wide$var_7TK <- as.numeric(naturetypes_wide$var_7TK)
```

```{r}
head(naturetypes_wide)
```

```{r}
anyDuplicated(naturetypes_wide$identifikasjon_lokalid)
```

@@

```{r, warning=F}
naturetypes_wide$slitasje <- apply(naturetypes_wide[,c("var_7TK", "var_7SE")], 1, max, na.rm=T)
```

When both variables are NA, we get -Inf. There were nine cases of this
(in 2023). Removing these now.

@@

ggplot(temp, aes(x = factor(slitasje),
                 y = sum))+
  geom_bar(stat="identity",
           fill="grey",
           colour = "black")+
  theme_bw(base_size = 12)+
  labs(x = "slitasje score",
       y = "Number of localities")
```

@@

I would also like to know how often 7TK was defining of the
slitasje-indicator, and how often it was 7SE. I can do this by taking
the difference

```{r, fig.cap="Counting the number of cases where 7TK or 7SE ws defining of the slitasje indicator score.", fig.width=3}
diff <- naturetypes_wide$var_7SE - 
        naturetypes_wide$var_7TK

>>>>

```{r}
myVars <- c('7TK', '7SE', 'PRTK', 'PRSL')
```

```{r}
naturetypes_summary <- naturetypes_summary_import %>%
  rowwise() %>%
  mutate(keepers = sum(c_across(
    myVars)>0, na.rm=T)) %>%
  filter(keepers > 0) %>%
  select(Nature_type, NiN_mainType, Year)
```

@@

Importing and sub-setting the main data file, fix duplicate _hovedøkosystem_, calculate area, split one column in two, make numeric, and select target variables:
```{r, message=F}
naturetypes <- sf::st_read(dsn = path) %>%
  filter(naturtype %in% naturetypes_summary$Nature_type) %>%
  mutate(hovedøkosystem = recode(hovedøkosystem,
                                 "naturligÅpneOmråderUnderSkoggrensa" = "naturligÅpneOmråderILavlandet"),
         area = st_area(.)) %>%
  separate_rows(ninBeskrivelsesvariable, sep=",") %>%
  separate(col = ninBeskrivelsesvariable,
           into = c("NiN_variable_code", "NiN_variable_value"),
           sep = "_",
           remove=F) %>%
  mutate(NiN_variable_value = as.numeric(NiN_variable_value)) %>%
  filter(NiN_variable_code %in% myVars)
```

@@

Now I need to create a single row per locality with a new variable which
is a product/function of the four variables "7SE", "7TK", "PRTK" and
"PRSL". I want to base this indicator on whichever of the variables have the highest value (worst condition). Therefore I first convert the ordinal scale to a continuous scale, using the median value of each ordinal step.

@@

```{r}
naturetypes <- naturetypes %>%
  mutate(NiN_variable_value = case_when(
    NiN_variable_code %in% c("var_7TK", "var_7SE") ~ 
      case_match(NiN_variable_value,
             0 ~ 0,
             1 ~ mean(c(0, 1/16))*100,
             2 ~ mean(c(1/16, 1/2))*100,
             3 ~ 75),
    NiN_variable_code %in% c('var_PRTK', 'var_PRSL') ~ 
      case_match(NiN_variable_value,
             0 ~ 0,
             1 ~ 1.5,
             2 ~ mean(c(3, 6.25)),
             3 ~ mean(c(6.25, 12.5)),
             4 ~ mean(c(12.5, 25)),
             5 ~ mean(c(25, 50)),
             6 ~ mean(c(50, 75)),
             7 ~ mean(c(75, 100)))
    )
  )
```

I will create a new data table where I calculate the new
variable which I then paste back into the sf object.

```{r}
naturetypes_wide <- pivot_wider(naturetypes,
                                names_from = "NiN_variable_code",
                                values_from = "NiN_variable_value",
                                id_cols = "identifikasjon_lokalId")
naturetypes_wide <- as.data.frame(naturetypes_wide)
head(naturetypes_wide, 10)
```

@@

```{r, warning=F}
naturetypes_wide$slitasje <- apply(naturetypes_wide[,-1], 1, max, na.rm=T)
```

When both variables are NA, we get -Inf. There were 13 cases of this
(in 2023). Removing these now.

@@

ggplot(temp, aes(x = factor(slitasje),
                 y = sum))+
  geom_bar(stat="identity",
           fill="grey",
           colour = "black")+
  theme_bw(base_size = 12)+
  labs(x = "Slitasje score converted to percentage",
       y = "Number of localities")+
  coord_flip()
```

@@

I would also like to know how often 7TK was defining of the
slitasje-indicator, and how often it was 7SE. I can do this by taking
the difference.

```{r, fig.cap="Counting the number of cases where 7TK or 7SE was defining of the slitasje indicator score.", fig.width=3}

diff <- max(naturetypes_wide$var_7SE, naturetypes_wide$var_PRSL, na.rm=T) - 
        max(naturetypes_wide$var_7TK, naturetypes_wide$var_PRTK, na.rm=T)
